---
title: CPAWS analysis for British Columbia Marine Protected Network under Climate
  Change
author: "Colette Wabnitz, Juliano Palacios Abrantes, William W.L. Cheung"
output:
  html_document:
    df_print: paged
---


```{r setup, message = F, eval = T, echo = F}
library(MyFunctions)

my_lib(
  c(
    "readxl",
    "tidyverse",
    "sf",
    "st",
    "janitor",
    # For grid estimation 
    "spatialEco", "geosphere","raster","units","matrixStats","rmapshaper", "igraph"
  )
)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select
```


# Pre-Analysis

# NOTES TO DISCUSS

- Observation, this shapefile has three dimensions, I am guessing there are some areas that are only bottom and this is why?
- Removed the depth component

There are (at least) three ways to reallocate fishing effort as it is in the DBEM. We are using area as follows"

- *Based on area*: To reallocate fishing from protected, we take the area protected/total area of grid cells that contain a protected area and reallocate that protected area proportion to the surrounding cells. This is the most accurate and least conservative (i.e least amount of fishing)

(area covered by mpa*protection level/total area of grid cells that contain a protected area)/n surrounding grid cells + 1 (or 1.5)

## Species explortion

# Load Species data

```{r species_list, eval = F, echo = F}

# NSB species list provided by Colette
nsb_species <- my_path("D","Species","updated_species.csv", read = T) 

# DBEM species
dbem_species <- my_path("Spp",name ="dbem_spp_list.csv", read = T)

# Exploited species
# exploited_species <- my_data("exploited_species") %>% 
# filter(taxon_key %in% dbem_species$taxon_key)

nsb_matching_species <- nsb_species %>% 
  filter(correct_name %in% dbem_species$worms_name) %>% 
  left_join(dbem_species %>% select(taxon_key,aphia_id, correct_name = worms_name)) %>% 
  write_csv(my_path("D","Species",name ="new_nsb_dbem_matching_spp.csv"))

# Update taxon list with new matching data

taxon_list <- my_path("D", "species", "CpawSppList10.txt", read = T, header = F) %>% 
  rename(taxon_key = v1)

missing_taxa <- nsb_matching_species %>% 
  filter(!taxon_key %in% taxon_list$taxon_key) %>% 
  select(taxon_key)

# Create species list for missing taxa
new_spp_list <- my_path("D","species","CpawSppList11.txt")
write.table(missing_taxa, file=new_spp_list, sep="\t", col.names = F, row.names = F)


```


## Spatial analysis

### Shapefile Data


```{r mpan_shape, eval = T, echo = F, message = F,warning = FALSE, cache = FALSE}

suppressMessages(
  mpan_sf <- st_read(my_path("D","Spatial/draft_mpan","DRAFT_MAP_NETWORK_NAP_P3_IOAC_Sept2022.shp")) %>% 
    # Remove the Z component for ease computation
    st_zm(drop = TRUE)
)
#get out coordinate reference system information 
# st_crs(mpan_sf)
st_crs(mpan_sf)$epsg
crs <- st_crs(mpan_sf)

mpan_sf_2d <- ggplot(mpan_sf) +
  geom_sf(aes(fill = as.character(UID_1))) +
  theme(legend.position = "")

# ggsave(filename = "../Results/Figures/shapefile_2d.png",mpan_sf_2d)
print(mpan_sf_2d)
```



# Grid shapefile

## Identify grids that are MPAs, surrounding an MPA and no-protected

## 1.1 Load in grid data

```{r grid, eval = T, echo = F}

grid <- read_sf(my_path("D","Spatial/worldsq_ea/", name ="worldsq_EA.shp"))

grid <- grid %>% filter(Lat > 48 & Lat < 56 & Lon > -136 & Lon < -120)

grid <- st_transform(grid, crs = st_crs(crs)) %>% 
  filter(PWater > 0)

ggplot(grid) +
  geom_sf()
```

## 1.2 Identify grids that are MPAs

In this step we identify which grid cells have an MPA completely or partially covering it. 

```{r mpa_overla, eval = T, echo = F, message = F}

# Intersect the grid and the MPA shapefile
grid_intersection <- st_intersection(grid, mpan_sf) %>%
  group_by(Seq) %>%
  summarise(geom = st_union(geometry)) %>%
  mutate(geom = st_sfc(geom),
         area = st_area(geom)) %>% 
  mutate(status = "protected") %>% 
  as.data.frame()

# Identify grid cells that have some level of MPA within them 
grid2 <- as.data.frame(grid) %>% 
  left_join(grid_intersection, by = "Seq") %>% 
  mutate(protected_area_m2 = as.numeric(area)) %>% 
  select(-geom) %>% 
  st_as_sf()

# Plot it
ggplot() + 
  geom_sf(data = grid2, 
          aes(fill = protected_area_m2)) +
  geom_sf(data = mpan_sf, 
          aes(), fill = NA, color = "white") +
  scale_fill_viridis_b()


```

## 1.3 Estimate area

Here we estimate the proportion of the grid cell that is protected using the area protected ($area_p$) relative to the area of the grid cell ($area_g$) in order to reduce fishing effort in grid celss that contain an MPA ($FE_{mpa}$):

$FE_{mpa} = \frac{area_p}{area_g}$

```{r area_est, eval = T, echo = F}

grid2 <- grid2 %>% 
  mutate(grid_area = st_area(.)) %>%
  drop_units() %>% 
  mutate(
    prop = ifelse(status == "protected", (protected_area_m2/grid_area), 1)
  ) 


ggplot() +
  geom_sf(data = grid2, aes(fill = prop)) +
  scale_fill_viridis_b("Proportion of grid\nprotected")
```

## 1.4 Find surrounding grid cells 

Surrounding grid cells are limited to the grid cells immediately adjacent to a grid cell that is cover (complete or partially) by an MPA. The identification of surrounding grid cells is necessary to further re-allocate fishing effort to these cells as they benefit from the biomass spillover.


```{r surroundig, eval = T, echo = F, message = F}

grid_mpan <- grid2 %>% 
  select(OBJECTID, Seq, Lat, Lon, geometry, status, protected_area_m2, grid_area)

# Identify non-protected
grid_mpan_no <- grid_mpan %>% filter(is.na(status))

#create buffer
buffer <- st_buffer(grid_mpan %>% filter(status == "protected"), 25000)

intr <- st_intersection(grid_mpan_no, buffer)
intr_seqs <- c(intr$Seq)

grid_surround <- grid_mpan_no %>% 
  filter(Seq %in% intr_seqs) %>% 
  mutate(surrounding ="surrounding") %>% 
  select(Seq, surrounding)


ggplot() +
  geom_sf(data = grid_surround , aes(), fill = "darkblue") +
  geom_sf(data = grid_mpan %>% filter(status == "protected"), aes(), fill = "pink") +
  geom_sf(data = mpan_sf, aes(), fill = NA, color = "black") +
  ggtitle("Showing protected grid cells in red and surrounding in blue")

```


# Reallocate fishing effort

Here we re-allocate fishing effort from the protected grid cells to those surrounding. We do this proportionally to the protected area and the number of surrounding grid cells. Right now, the proportion is accounted from all MPAs and re-distributed to all grid cells, but this can change. For this we first estimate the proportion to be allocated (*PropAllocate*):


$$PropAllocate = \frac{tot_{mpa}}{tot_{area}}$$

where $tot_{mpa}$ is the sum of the protected area ($m^2$) of all protected grid cells and $tot_{area}$ is the sum of the whole grided area (protected and surrounding). Then, we estimate the allocation of effort by first counting the number of surrounding grid cells and then allocating the proportion to the fishing effort of surrounding cells ($FE_{sur}$):

$$FE_{sur} = 1+\frac{PropAllocate}{n_{surround}}$$
This way, fishing effort in an MPA grid will be reduced by the proportion of the MPA in the grid ($FE_{mpa}$), in surroundings cells will be increased with the proportion allocated ($FE_{sur}$) while all other grid cells (no-protected) will have an DF at MSY.

```{r fiseffort, eval = T, echo = F}

#total proportion 
#this will help allocate fishing effort to the surrounding areas. 
tot_area = sum(grid_mpan$grid_area, na.rm = T)
tot_mpa = sum(grid_mpan$protected_area_m2, na.rm = T)
prop = tot_mpa/tot_area


grid_use <- left_join(grid2, grid_surround %>% st_drop_geometry(), by = "Seq")

grid_sc1 <- grid_use %>% 
  mutate(prop = if_else(status %in% "protected", 1-prop, prop)) #these should all be 0

#grid cells surrounding mpas
n_surround <- grid_sc1 %>% 
  st_drop_geometry() %>% 
  filter(surrounding == "surrounding") %>% 
  tally()

#protected proportion to reallocate (calculated as area of each mpa/area of grid cells containing mpa)
prop_allocate = tot_mpa/tot_area

#allocate surrounding cells
grid_sc1 <- grid_sc1 %>%  
  mutate(FE_prop = if_else(surrounding %in% "surrounding",(prop_allocate/n_surround$n) +1 , prop),
         status = ifelse(surrounding %in% "surrounding","surrounding",status),
         status = ifelse(is.na(status),"unprotected",status)
  )

ggplot() +
  geom_sf(data = subset(grid_sc1, grid_sc1$FE_prop != 1),
          aes(fill = FE_prop)) +
  geom_sf(data = mpan_sf, aes(), fill = NA, color = "white") +
  scale_fill_viridis_b()

```

## Incorporating grid into DBEM

```{r}

# Get grid from previous steps
mpa_grid <- as.data.frame(grid_sc1) %>% 
  ungroup() %>% 
  select(index = OBJECTID,
         lon = Lon,
         lat = Lat,
         fe_prop = FE_prop)

# Load DBEM grid
dbem_grid <- my_data("dbem_coords") %>% 
  filter(!index %in% mpa_grid$index) %>%  #filter out BC
  bind_rows(mpa_grid) %>% 
  mutate(fe_prop = ifelse(is.na(fe_prop), 1,fe_prop)) #include fishing proportion

# Map it to make sure it works
dbem_grid %>% 
  filter(lat > 40,
         lon < -100) %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = log10(fe_prop)
    )
  ) +
  scale_fill_viridis_b()


# Prepare it to the DBEM
dbem_grid_index <- dbem_grid %>% 
  arrange(index) %>% 
  select(fe_prop)

# Create index
name <- my_path("D","spatial","mpa_cpaw.txt")
write.table(dbem_grid_index, file=name, sep="\t", col.names = F, row.names = F)


```

# Species selection

```{r}
# Load updated species list by Colette
# Note these species have been crossreff. with the WORMS database for taxonomy
nsb_species <- my_path("D","species", "nsb_updated_species.csv", read = T)

# Load DBEM species list
dbem_species <- my_data("dbem_species") %>% 
  rename(scientific_name = worms_name)
  
# Save spp list for DBEM (n = 45)
CpawSppList <- nsb_species %>% 
  left_join(dbem_species,
            by = "scientific_name") %>% 
  select(taxon_key) %>% 
  filter(!is.na(taxon_key))

# Create index
name_spp_list <- my_path("D","species","CpawSppList12.txt")
write.table(CpawSppList, file=name_spp_list, sep="\t", col.names = F, row.names = F)

# Note CpawSppList12 is the most recent species list to run ONLY with species data

```

## Genus matching

There are some species of interest that do not have species data information to run the DBEM with. For these species we will use an average of genus


```{r}

# Match species with missing data with genus
missing_s <- my_path("D","species", "nsb_updated_species.csv", read = T) %>%
  select(scientific_name, genus_other) %>% 
  left_join(dbem_species,
            by = c("scientific_name")
  ) %>% 
  filter(is.na(taxon_key)) %>% 
  select(scientific_name, genus = genus_other) %>% 
  left_join(my_data("dbem_species"),
            by = "genus") %>% 
  select(scientific_name,genus,taxon_sub = taxon_name,taxon_key)

# Save data
write_csv(
  missing_s %>% filter(!is.na(taxon_key)),
  my_path("D","Species",name ="species_to_genus_sub.csv", read = F)
)


# Get species that will use substitute genus data (n = 34)
species_to_sub  <- missing_s %>% 
  filter(!is.na(taxon_key)) %>% 
  group_by(scientific_name) %>% 
  tally() 

# Still some species missing data even at the genus level (n = 63)
missing_g <- missing_s %>% 
  filter(is.na(taxon_key)) %>% 
  group_by(scientific_name) %>% 
  tally() 


# Full list of species and their data sources
spp_data_source <- my_path("D","Species",name ="nsb_updated_species.csv", read = T) %>%
  mutate(
    data_source = ifelse(scientific_name %in% species_to_sub$scientific_name, "genus",
                         ifelse(scientific_name %in% missing_g$scientific_name, "no data",
                                "species"
                         )
    )
  )

# Save species data
write_csv(
  spp_data_source,
  my_path("D","Species",name ="species_data_sources.csv", read = F)
)

```



