---
title: "Projection analysis"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_notebook:
    code_folding: hide
    number_sections: true
    theme: cerulean
    highlight: haddock
---

```{r setup, message = F, eval = T, echo = F}

library(MyFunctions)

my_lib(
  c(
    "tidyverse",
    "sf",
    "st",
    "janitor",
    "zoo",
    "knitr",
    "kableExtra"
  )
)

# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select
```

# Overall

This script analyzes 4 scenarios developed for the CPAWS projec on climate change in the Coast of British Columbia. The 4 scenarios were created in the script `initial_analysis_nb.Rmd` and they all reduce fishing mortality in the protected grid cells by 25%, 50%, 75% and 100% of $f/f_{msy}$ proportionally to the area of the grid cell that is covered by the MPA. This means that, for the 50% scenario, if an MPA covers 25% of a given protected grid cell, then, the actual reduction would be 12.5% (0.50-scenario × 0.25-area cover = 0.125 or 12.5%), instead of 50%.
# Load data

```{r load_data, eval = T}

# Get modelled species DF
# taxon_list <- my_path("D","Species",name ="nsb_updated_species.csv", read = T) %>% 
#   select(taxon_name = scientific_name) %>% 
#   left_join(my_data("dbem_species")) %>% 
#   filter(!is.na(taxon_key)) # remove species not modeled


ran_taxa <- list.files("/Volumes/Enterprise/Data/cpaws_bc/Data/dbem_runs/txt/c6gfdl85f1cpaws25/")

taxon_list <- my_data("dbem_species") %>% 
  filter(taxon_key %in% ran_taxa)
  
  

# manually include herring nor dungeness crab
dbem_species <- my_data("dbem_species") %>% 
  filter(taxon_key %in% c(690115,601520))

# 
# taxon_list <- taxon_list %>% 
#   bind_rows(dbem_species)


# Get MPA sf
suppressMessages(
  bs_sf <- st_read(my_path("D","Spatial/draft_mpan","DRAFT_MAP_NETWORK_NAP_P3_IOAC_Sept2022.shp")) %>% 
    # Remove the Z component for ease computation
    st_zm(drop = TRUE) %>% 
    st_simplify(preserveTopology = TRUE, dTolerance = 100)
) 
st_crs(bs_sf)$epsg
crs <- st_crs(bs_sf)

bs_sf <- st_transform(bs_sf,4326)

# Get EEZ
bc_eez <- my_sf("SAU") %>% 
  filter(eezid == 925)

# Get BC grid to filter out
bc_grid <- my_data("sau_index") %>% 
  filter(eez_name == "Canada (Pacific)") %>% 
  pull(index)

```



# Plot maps per species


```{r eval = F, create_maps}


# FOR NON MPA RUNS
# Get DBEM paths
# dbem_no_mpa <- my_path("G",# extra_path = "raw/C6GFDL26F1CPAW", #SSP126 for now
#                        extra_path = "dbem/dbem_cmip6/r_data/C6GFDL26F1_t2", #SSP126 for now
#                        list_files = "paths"
                       # )

# For non-mpa runs
# Filter out paths from speceis that we are interested in
# dbem_no_mpa <- dbem_no_mpa[grep(paste(taxon_list$taxon_key, collapse = "|"), dbem_no_mpa)]

# lapply(dbem_no_mpa, pr_chg_fx, mpa = F) # Multiple



# For MPA runs
scenarios <- c(25,50,75,10)
for(i in 1:3){
  
  
dbem_mpa_data <- my_path("D",
                         extra_path = paste0("dbem_runs/rdata/c6gfdl85f1cpaws",scenarios[i]),
                         list_files = "paths",
                         fn = T,
                         system = "juliano")


dbem_mpa_data <- dbem_mpa_data[grep(paste(taxon_list$taxon_key, collapse = "|"), dbem_mpa_data)]



lapply(dbem_mpa_data, pr_chg_fx, mpa = T) # Multiple

}


```


# MPA differences per species

```{r}

to_read <- paste0("/Volumes/Enterprise/Data/cpaws_bc/Data/dbem_runs/rdata/c6gfdl85f1cpaws",scenarios,"/","600090Abd.RData")

test_path <- to_read[1]

scen_test <- function(test_path){
  load(test_path)
  #Global variables
  esm = str_sub(test_path,58,61)
  ssp = ifelse(str_sub(test_path,62,63) == 26,"ssp126","ssp585")
  scen = ifelse(str_sub(test_path,71,72) == 10,"0",paste0(str_sub(test_path,71,72)))
  
  # Load Abd or MCP data
  
  data <- as.data.frame(data) %>% 
    rowid_to_column("index")
  colnames(data) <- c("index",(seq(1851,2100,1)))
  if(str_detect(test_path,"Abd") == T){
    variable <- "abd"
  }else{
    variable <- "mcp"
  }
  
  data <- data %>% 
    select(index,`1951`:`2100`) %>% 
    filter(index %in% bc_grid) %>% 
    gather("year","value",`1951`:`2100`) %>% 
    mutate(
      variable = variable,
      esm=esm,
      ssp = ssp,
      scenario = scen
    )
  
  return(data)
  
}


scen_test_df <- bind_rows(
  lapply(to_read, scen_test)
)

test_map <- scen_test_df %>% 
  left_join(dbem_grid) %>% 
  filter(!is.na(value)) %>% 
  ggplot() +
  geom_tile(
        aes(
          x = lon,
          y = lat,
          fill = value,
          color = value
        )
      ) +
      scale_fill_viridis_c("Abundance") +
      scale_color_viridis_c("Abundance") +
      # extra maps
      geom_sf(data = bs_sf, aes(), color = "black", fill = "transparent") +
      geom_sf(data = bc_eez, aes(), color = "grey50", fill = "transparent") +
      MyFunctions::my_land_map() +
      labs(
        x = "Longitude",
        y = "Latitude"
      ) +
      coord_sf(
        xlim = c(-120,-140),
        ylim = c(46,56)
      ) +
      theme_classic()+
      facet_wrap(~scenario)


ggsave("test_map.png",
       test_map)

scen_test_df %>% 
  group_by(year,variable,esm,ssp,scenario) %>% 
  summarise(
    value_sum = sum(value,na.rm = T)
  ) %>% 
  ggplot() +
  geom_line(
        aes(
          x = as.numeric(year),
          y = value_sum,
          color = scenario
        )
      ) +
      scale_fill_viridis_d("scenario") +
      scale_color_viridis_d("scenario") +
      theme_classic()



```



# Lineal plot of projection differences 

## Per EEZ

### Abundance

These plots represent the total abundance in the whole BC EEZ for each species, regardless of the "conservation status" of the grid cell. We basically add up all gridcells in the EEZ by species, year and DBEM run (i.e., run considering the MPA grid and runs without). Remember that absolute DBEM values should be seen as reference points, and not real estimates of population abundance in the water.

Overall, it is expected that the MPA runs will have more abundance as fishing is closed in some areas (e.g., grid cells that cointain an MPA) so naturally, there is less $f_{mort}$ happening within the EEZ.

```{r abd_summary, eval = F}



agg_all_scen <- function(scenario,ssp,grid){
  
  
  dbem_paths <- taxon_list %>% 
    filter(taxon_key %in% ran_taxa) %>% 
    distinct() %>% 
    select(taxon_key) %>% 
    mutate(
      dbem_no_mpa_path = paste0("~/Library/CloudStorage/OneDrive-UBC/Data/dbem/dbem_cmip6/r_data/C6GFDL",ssp,"F1_t2/",taxon_key,variable,".RData"),
      dbem_mpa_path = paste0("/Volumes/Enterprise/Data/cpaws_bc/Data/dbem_runs/rdata/c6gfdl",ssp,"f1cpaws",scenario,"/",taxon_key,variable,".RData")
    )
  
  
  # Call function
  
  agg_per_species_abd <-agg_per_spp_abd(dbem_paths, grid = grid)
  
  return(agg_per_species_abd)
  
}



variable = "Abd"

# For MPA runs
scenarios <- c(25,50,75,10)

agg_per_species_abd_scen <- bind_rows(
  lapply(scenarios, agg_all_scen, ssp = 85, grid = F)
  
)


unique(agg_per_species_abd_scen$scenario)
unique(agg_per_species_abd_scen$ssp)
unique(agg_per_species_abd_scen$variable)
unique(agg_per_species_abd_scen$run)
length(unique(agg_per_species_abd_scen$taxon_key))

write_csv(agg_per_species_abd_scen, my_path("R","tables","species_abd_gfdl85.csv"))

```

#### Total abundance

Figure showing the total abundance (raw DBEM outputs) per species over the years across the whole BC. EEZ. 

```{r total_abd, eval = T, fig.width=20, fig.height=15, message=F,warning=F}

# read data
agg_per_species_abd <- my_path("R","tables","species_abd_gfdl85.csv",read = T) %>% 
  mutate(scenario = ifelse(scenario == 100,0,scenario)) %>% 
  left_join(taxon_list,
            by = ("taxon_key")
            ) %>% 
  filter(taxon_key != 600309) #greenland cod not working

# Raw data
plot_agg_per_species_abd <- agg_per_species_abd %>% 
  left_join(taxon_list) %>% 
  filter(run == "mpa") %>% 
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = sum,
      col = as.character(scenario)
      # linetype = run
    )
  ) +
  facet_wrap(~common_name, scales = "free") +
  theme_classic()+
  theme(
  text = element_text(size = 20)  # Set base text size
)

ggsave(my_path("R","figures","spp_abd_gfdl85.png"),
       plot = plot_agg_per_species_abd,
       width = 20,
       height = 15)


# print(plot_agg_per_species_abd)
```

#### Difference in abundance 

##### Relative to the no mpa run

Figure showing the difference in abundance between the MPA and the no-MPA runs per species over the years across the whole BC. EEZ. Percentage difference in abundance ($\Delta{abd}$) was calculated as:

$$ \Delta{abd} = \frac{mpa-nompa}{nompa} * 100$$

Thus, results should be read as; positive values representing more abundance in the MPA runs and negative values more abundance in the non-MPA runs. Note that $ \Delta{abd}$ was caped at 100%/-100% for visualization purposes. Also showing a 10 years running mean. One can argue that the larger the difference, the larger the effect of MPA in the population.

**Note:** This type of result has been contested previously on peer review as "logical" due to the nature of the model.


```{r diff_abd, eval = T, fig.width=20, fig.height=15, message=F, warning = F}

# Raw data
diff_plot <-
  agg_per_species_abd %>% 
  left_join(taxon_list) %>% 
  select(-variable,-mean) %>% 
  group_by(year, taxon_key,common_name,scenario,esm,ssp) %>%
  summarize(diff = sum[run == "mpa"] - sum[run == "no_mpa"],
            percent_diff = (diff / sum[run == "no_mpa"]) * 100
  ) %>% 
  mutate(percent_diff_capped = pmax(pmin(percent_diff, 100), -100)) %>% 
  group_by(taxon_key,common_name,scenario,esm,ssp) %>%
  arrange(year) %>%
  mutate(
    run_avg = rollmean(percent_diff, k = 10, align = "right", fill = NA)
  ) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = percent_diff_capped,
      color = as.character(scenario)
    )
  ) +
  geom_line(
    aes(
      x = as.numeric(year),
      y = run_avg,
      # group = taxon_key,
      color = as.character(scenario)
    )
  ) +
  theme_classic()+
  theme(
  text = element_text(size = 20)  # Set base text size
  ) +
  facet_wrap(~common_name, scales = "free")


ggsave(my_path("R","figures","spp_abs_diff_per_scen_gfdl26.png"),
       plot = diff_plot,
       width = 20,
       height = 15)

print(diff_plot)
```


##### Relative to the 100 scenario

```{r diffabdscenarios, eval = T, fig.width=20, fig.height=15, message=F, warning = F}

# Raw data
diff_plot <-
  agg_per_species_abd %>% 
  filter(run == "mpa") %>% 
  dplyr::select(-variable,-mean) %>%
  # dplyr::select(year,scenario,taxon_key,sum) %>% 
  # View()
  pivot_wider(names_from = scenario, values_from = sum) %>% 
  mutate(
    d_25 = my_chng(`0`,`25`),
    d_50 = my_chng(`0`,`50`),
    d_75 = my_chng(`0`,`75`)
  ) %>% 
  gather("delta","delta_value",d_25:d_75) %>% 
  group_by(taxon_key,common_name,esm,ssp,delta) %>%
  # arrange(year)
  mutate(
    delta_value = ifelse(delta_value > 100,100,
                         ifelse(delta_value < -100,-100,delta_value)),
    run_avg = rollmean(delta_value, k = 10, align = "right", fill = NA)
  ) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = delta_value,
      color = as.character(delta)
    )
  ) +
  theme_classic()+
  theme(
  text = element_text(size = 20)  # Set base text size
  ) +
  facet_wrap(~common_name, scales = "free")


ggsave(my_path("R","figures","spp_abs_diff_scen_gfdl85.png"),
       plot = diff_plot,
       width = 20,
       height = 15)


```



#### Relative difference

The following figure and table shows the percentage difference between the MPA and the no-MPA runs by an early period (2050) and the end of the century (2100). For now, periods are defined as:

- "historic", year > 1995 & year < 2014, 
- "early_2050", year > 2041 & year < 2060
- "end_2100", year > 2081 & year < 2100


Here, the difference (`diff_early`) represents the difference in percentage change of the two models. Taking Capelin as an example, in the MPA runs, capelin has a historically abundance of 120,121 tones which decreases by 55% (to 53,624 tones) by 2050 and 60% (to 45,769 tones) by 2100. In the no-MPA runs, the numbers are slightly different with a historical abundance of 102,222, a decrease of 54% (to 46,829 tons) by 2050 and a total decrease of 60% (to 40,382) by 2100. That means that, for Capelin, the next figure shows `diff-early = -1%` (-55 - -54) and a `diff-end = -2%`.

*Note: The most common reason why the MPA model often presents higher loss of biomass is because it has higher overall biomass, so the loss is larger, nevertheless, the absolute biomass is still higher in the MPA model than in the non-MPA model (as shown in figure 1).*


```{r relative_abd_tbl, eval = T, fig.width=8, fig.height=7, message=F,warning=F}
# Raw data

abd_relative_df <- agg_per_species_abd %>% 
  select(-variable,-mean) %>% 
  mutate(
    time_frame = ifelse(year > 1995 & year < 2014, "historic",
                        ifelse(year > 2041 & year < 2060, "early_2050",
                               ifelse(year > 2081 & year < 2100, "end_2100",NA)
                        )
    )
  ) %>% 
  filter(!is.na(time_frame)) %>% 
  group_by(run,time_frame, taxon_key,common_name,esm,ssp,scenario) %>%
  summarise(nat_mean_sum = round(mean(sum,na.rm = T))) %>% 
  group_by(taxon_key,common_name,esm,ssp,scenario) %>% 
  spread(time_frame,nat_mean_sum) %>% 
  mutate(early_2050 = my_chng(`historic`,`early_2050`,limit = 100),
         end_2100 = my_chng(`historic`,`end_2100`,limit = 100)
  )


print(abd_relative_df %>% select(1:3,historic,everything()) %>% filter(common_name == "Capelin"))

```


```{r}

abd_relative_df %>% 
  gather("time_frame","percentage_difference",early_2050:end_2100) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = percentage_difference,
      fill = as.character(scenario)
    ),
    stat = "identity",
    position = "dodge"
  ) +
  scale_fill_viridis_d("F scenario") +
  coord_flip() +
  labs(x = "Common name",
       y = "Percentage change relative to historical values") +
  theme_classic() +
  theme(
    text = element_text(size = 16)  # Set base text size
  ) +
  facet_wrap(~time_frame)


ggsave(my_path("R","figures","spp_time_diff_abd_gfdl85_scen.png"),
       plot = last_plot(),
       width = 10,
       height = 8)


```


##### Relative to the no mpa run

```{r relative_abd_fig_nompa, eval = T, fig.width=8, fig.height=7, message=F,warning=F}

rel_diff <- abd_relative_df %>% 
  group_by(taxon_key,common_name,esm,ssp,scenario) %>%
  # View()
  summarize(
    diff_early = per_chng_early[run == "mpa"] - per_chng_early[run == "no_mpa"],
    diff_end = per_chng_end[run == "mpa"] - per_chng_end[run == "no_mpa"]
  ) %>% 
  # View()
  gather("time_frame","percentage_difference",diff_early:diff_end)

rel_diff %>% 
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = percentage_difference,
      color = as.character(scenario),
      fill = as.character(scenario)
    ),
    stat = "identity",
    position = "dodge"
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    text = element_text(size = 16)  # Set base text size
  ) +
  facet_wrap(~time_frame)

ggsave(my_path("R","figures","rel_diff_per_species_abd_gfdl85_scen.png"),
       plot = last_plot(),
       width = 10,
       height = 8)

print(rel_diff)
  
```

##### Relative to the 100 scenario (Delta delta)

```{r relative_abd_fig_deltascen, eval = T, fig.width=8, fig.height=7, message=F,warning=F}

delta_delta_scen <-
  abd_relative_df %>% 
  filter(run == "mpa") %>% 
  select(-historic) %>% 
  gather("time_frame","percentage_difference",early_2050:end_2100) %>% 
  spread(scenario,percentage_difference) %>% 
  mutate(
    d_25 = my_chng(`0`,`25`),
    d_50 = my_chng(`0`,`50`),
    d_75 = my_chng(`0`,`75`)
  ) %>% 
  gather("delta","delta_value",d_25:d_75) %>% 
  mutate(
    delta_value = ifelse(delta_value > 100,100,
                         ifelse(delta_value < -100,-100,delta_value))
    
  )

delta_delta_scen %>% 
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = delta_value,
      fill = as.character(delta)
    ),
    stat = "identity",
    position = "dodge"
  ) +
  scale_fill_viridis_d("F scenario") +
  coord_flip() +
  labs(x = "Common name",
       y = "Difference in climate change impacts between F scenarios\nrelative to 100% protection (%)") +
  theme_classic() +
  theme(
    text = element_text(size = 16)  # Set base text size
  ) +
  facet_wrap(~time_frame)

ggsave(my_path("R","figures","spp_delta_delta_abd_scen_gfdl85.png"),
       plot = last_plot(),
       width = 10,
       height = 8)


  
```


### Aggregated

```{r}

agg_per_species_abd <- my_path("R","tables","species_abd_gfdl85.csv",read = T) %>% 
  mutate(scenario = ifelse(scenario == 100,0,scenario)) %>% 
  group_by(run,variable,year,scenario,esm,ssp) %>% 
  summarise(
    total_value = sum(sum,na.rm = T)
  )

```

#### Difference in abundance 

```{r}
agg_per_species_abd %>% 
  filter(run == "mpa") %>% 
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = total_value,
      col = as.character(scenario),
      # linetype = run
    )
  ) +
  theme_classic()+
  scale_color_viridis_d("F Scenario") +
  labs(
    x = "Year",
    y = "DBEM relative abundance"
  ) +
  theme(
  text = element_text(size = 20)  # Set base text size
)

ggsave(my_path("R","figures/aggregated","agg_abd_gfdl85.png"),
       plot = last_plot(),
       width = 10,
       height = 7)
```



```{r diffabdscenarios_totals, eval = T, fig.width=20, fig.height=15, message=F, warning = F}

# Raw data

  agg_per_species_abd %>% 
  filter(run == "mpa") %>% 
  spread(scenario,total_value) %>% 
  mutate(
    d_25 = my_chng(`0`,`25`,limit = 100),
    d_50 = my_chng(`0`,`50`,limit = 100),
    d_75 = my_chng(`0`,`75`,limit = 100)
  ) %>% 
  gather("delta","delta_value",d_25:d_75) %>% 
  group_by(esm,ssp,delta) %>%
  # arrange(year)
  mutate(
    run_avg = rollmean(delta_value, k = 10, align = "right", fill = NA)
  ) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = delta_value,
      color = as.character(delta)
    )
  ) +
  # geom_line(
  #   aes(
  #     x = as.numeric(year),
  #     y = run_avg,
  #     group = taxon_key,
  #     color = as.character(delta)
  #   )
  # ) +
  scale_color_viridis_d("F Scenario") +
  labs(
    x = "Year",
    y = "Percentage change relative to F0 scenario"
  ) +
  theme_classic()+
  theme(
  text = element_text(size = 20)  # Set base text size
  )


ggsave(my_path("R","figures/aggregated/","agg_abd_scen_diff_gfdl85b.png"),
       plot = last_plot(),
       width = 10,
       height = 7)


```

#### Relative difference

```{r relative_abd_tbl_agg, eval = T, fig.width=8, fig.height=7, message=F,warning=F}
# Raw data

abd_relative_df <- agg_per_species_abd %>% 
  mutate(
    time_frame = ifelse(year > 1995 & year < 2014, "historic",
                        ifelse(year > 2041 & year < 2060, "early_2050",
                               ifelse(year > 2081 & year < 2100, "end_2100",NA)
                        )
    )
  ) %>% 
  filter(!is.na(time_frame)) %>% 
  group_by(run,time_frame, esm,ssp,scenario) %>%
  summarise(nat_mean_sum = round(mean(total_value,na.rm = T))) %>% 
  group_by(esm,ssp,scenario) %>% 
  spread(time_frame,nat_mean_sum) %>% 
  mutate(early_2050 = my_chng(`historic`,`early_2050`,limit = 100),
         end_2100 = my_chng(`historic`,`end_2100`,limit = 100)
  )


print(abd_relative_df %>% select(1:3,historic,everything()) %>% filter(common_name == "Capelin"))

```

```{r}

abd_relative_df %>% 
  gather("time_frame","percentage_difference",early_2050:end_2100) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = time_frame,
      y = percentage_difference,
      fill = as.character(scenario)
    ),
    stat = "identity",
    position = "dodge"
  ) +
  scale_fill_viridis_d("F scenario") +
  # coord_flip() +
  labs(x = "Common name",
       y = "Percentage change relative to historical values") +
  theme_classic() +
  theme(
    text = element_text(size = 16)  # Set base text size
  ) 


ggsave(my_path("R","figures/aggregated","agg_abd_time_diff_gfdl85.png"),
       plot = last_plot(),
       width = 10,
       height = 8)


```

##### Relative to the 100 scenario (Delta delta)

```{r relative_abd_fig_total, eval = T, fig.width=8, fig.height=7, message=F,warning=F}

abd_relative_df %>% 
  filter(run == "mpa") %>% 
  select(-historic) %>% 
  gather("time_frame","percentage_difference",early_2050:end_2100) %>% 
  spread(scenario,percentage_difference) %>% 
  mutate(
    d_25 = my_chng(`0`,`25`),
    d_50 = my_chng(`0`,`50`),
    d_75 = my_chng(`0`,`75`)
  ) %>% 
  gather("delta","delta_value",d_25:d_75) %>% 
  # View()
  ggplot() +
  geom_bar(
    aes(
      x = delta,
      y = delta_value,
      fill = as.character(delta)
    ),
    stat = "identity",
    position = "dodge"
  ) +
  scale_fill_viridis_d("F scenario") +
  coord_flip() +
  labs(x = "Scenario",
       y = "Difference in climate change impacts between F scenarios\nrelative to 100% protection (%)") +
  theme_classic() +
  theme(
    text = element_text(size = 16)  # Set base text size
  ) +
  facet_wrap(~time_frame)

ggsave(my_path("R","figures/aggregated","agg_delta_delta_abd_scen_gfdl85.png"),
       plot = last_plot(),
       width = 8,
       height = 4)


  
```


### Maximum Catch Potential

```{r mcp_summary, eval = F}

variable = "Catch"

dbem_paths <- taxon_list %>% 
  distinct() %>% 
  select(taxon_key) %>% 
  mutate(dbem_no_mpa_path = paste0("~/Library/CloudStorage/OneDrive-UBC/Data/dbem/dbem_cmip6/r_data/C6GFDL26F1_t2/",taxon_key,"MCP.RData"),
         dbem_mpa_path = paste0("~/Library/CloudStorage/OneDrive-UBC/Data/cpaws_bc/Results/raw/C6GFDL26F1CPAW/",taxon_key,variable,".RData")
  )


agg_per_species_mcp <- agg_per_spp_mcp(dbem_paths)

write_csv(agg_per_species_mcp, my_path("R","tables","agg_per_species_mcp.csv"))

```

#### Total MCP

Figure showing the total MCP (raw DBEM outputs) per species over the years across the whole BC. EEZ. 

##### Facet figure

```{r mcp_total_facet, eval = T, echo = T, fig.width = 20, fig.height = 15, message=F,warning=F}

# read data
agg_per_species_mcp <- my_path("R","tables","agg_per_species_mcp.csv",read = T) %>% 
  left_join(taxon_list)

# Raw data
plot_agg_per_species_mcp<- ggplot(agg_per_species_mcp) +
  geom_line(
    aes(
      x = as.numeric(year),
      y = sum,
      col = run
    )
  ) +
  facet_wrap(~common_name, scales = "free") +
  theme_classic()+
  theme(
  text = element_text(size = 20)  # Set base text size
)


ggsave(my_path("R","figures","agg_per_species_mcp.png"),
       plot = plot_agg_per_species_mcp,
       width = 20,
       height = 15)


print(plot_agg_per_species_mcp)

```


##### Facet per group

```{r mcp_total_group, eval = T, echo = T, fig.width = 20, fig.height = 15, message=F,warning=F}

# read data
agg_per_species_mcp <- my_path("R","tables","agg_per_species_mcp.csv",read = T) %>% 
  left_join(taxon_list)

# Taxa where MPA and no MPA runs are the same
same_taxa <- c(600142,600243,600252,600255,600256,600502,601477,604009,604238)

# Groups
average_sepration <- agg_per_species_mcp %>% 
  group_by(run,taxon_key,taxon_name,common_name) %>% 
  summarise(
    mean_year = mean(sum,na.rm = T)
  ) %>% 
  spread(run,mean_year) %>% 
  mutate(
    flag = ifelse(mpa > no_mpa, "MPA",
                  ifelse(taxon_key %in% same_taxa,"No difference",
                         "No MPA")
    )
  ) %>% 
  select(taxon_key,flag) %>% 
  filter(taxon_key != 601732)


# Raw data

for(i in 1:3){
  
  value <- c("MPA","No MPA","No difference")[i]

plot_agg_per_species_mcp <- agg_per_species_mcp %>% 
  left_join(average_sepration) %>% 
  filter(flag == value) %>% 
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = sum,
      col = run
    )
  ) +
  facet_wrap(~common_name, scales = "free") +
  theme_classic()+
  labs(
    x = "Year",
    y = "Total MCP within study region"
  )
  theme(
  text = element_text(size = 10)  # Set base text size
) + 
  ggtitle(paste0("Species where the average MCP across all years is higher with ",value))


ggsave(my_path("R","figures",paste0("aagg",value,"_per_species_mcp.png")),
       plot = plot_agg_per_species_mcp,
       width = 10,
       height = 8)

}



# 601732 (remove)

```


##### Per species figure

```{r mcp_total_per_spp, eval = T, echo = T, fig.width = 20, fig.height = 15, message=F,warning=F}

# read data
agg_per_species_mcp <- my_path("R","tables","agg_per_species_mcp.csv",read = T) %>% 
  left_join(taxon_list)




for(i in 1:length(unique(taxon_list$taxon_name))){
  
  taxon <- unique(taxon_list$taxon_name)[i]
  taxon_name <- unique(taxon_list$common_name)[i]
  
  plot_agg_per_species_mcp <- agg_per_species_mcp %>% 
    filter(taxon_name == taxon) %>% 
    ggplot() +
    geom_line(
      aes(
        x = as.numeric(year),
        y = sum,
        col = run
      )
    ) +
    facet_wrap(~common_name, scales = "free") +
    theme_classic()+
    labs(x = "Year",
         y = "Total MCP") +
    theme(
      text = element_text(size = 10)  # Set base text size
    ) +
    ggtitle(paste0("Projected MCP for ",taxon))
  
  
  png_name <- paste0(taxon_name,".png")
  
  ggsave(my_path("R","species_mcp",png_name),
       plot = plot_agg_per_species_mcp,
       width = 7,
       height = 5)



}




```


#### Difference in MCP

```{r mcp_diff, eval = T, echo = T, fig.width = 15, fig.height = 10, message=F,warning=F}

# Raw data
diff_plot_mcp<- agg_per_species_mcp %>% 
  select(-variable,-mean) %>% 
  group_by(year, taxon_key,common_name) %>%
  summarize(diff = sum[run == "mpa"] - sum[run == "no_mpa"],
            percent_diff = (diff / sum[run == "no_mpa"]) * 100
  ) %>% 
  mutate(percent_diff_capped = pmax(pmin(percent_diff, 100), -100)) %>% 
  group_by(taxon_key,common_name) %>%
  arrange(year) %>%
  mutate(
    run_avg = rollmean(percent_diff, k = 10, align = "right", fill = NA)
  ) %>% 
  # View()
  ggplot() +
  geom_line(
    aes(
      x = as.numeric(year),
      y = percent_diff_capped,
      color = "percent_diff_capped"
    )
  ) +
  geom_line(
    aes(
      x = as.numeric(year),
      y = run_avg,
      group = taxon_key,
      color = "running_mean"
    )
  ) +
  facet_wrap(~common_name, scales = "free") +
  theme_classic()+
  theme(
    text = element_text(size = 20)  # Set base text size
  )



ggsave(my_path("R","figures","abs_diff_per_species_mcp.png"),
       plot = diff_plot_mcp,
       width = 20,
       height = 15)

print(diff_plot_mcp)

```

#### Relative difference


```{r mcp_relative_diff, eval = T, echo = T, fig.width = 8, fig.height = 5, message=F,warning=F}

# Raw data
mcp_relative_df <- agg_per_species_mcp %>% 
  select(-variable,-mean) %>% 
  mutate(
    time_frame = ifelse(year > 1995 & year < 2014, "historic",
                        ifelse(year > 2041 & year < 2060, "early_2050",
                               ifelse(year > 2081 & year < 2100, "end_2100",NA)
                        )
    )
  ) %>% 
  filter(!is.na(time_frame)) %>% 
  group_by(run,time_frame, taxon_key,common_name) %>%
  summarise(nat_mean_sum = round(mean(sum,na.rm = T))) %>% 
  group_by(taxon_key,common_name) %>% 
  spread(time_frame,nat_mean_sum) %>% 
  mutate(per_chng_early = my_chng(`historic`,`early_2050`,limit = 100),
         per_chng_end = my_chng(`historic`,`end_2100`,limit = 100)
  ) %>% 
  arrange(common_name)

print(mcp_relative_df %>% select(1:3,historic,everything()) %>% filter(common_name == "Capelin"))

```


```{r}



rel_diff_mcp<- mcp_relative_df %>% 
  group_by(taxon_key,common_name) %>%
  # View()
  summarize(
    diff_early = per_chng_early[run == "mpa"] - per_chng_early[run == "no_mpa"],
    diff_end = per_chng_end[run == "mpa"] - per_chng_end[run == "no_mpa"]
  ) %>% 
  gather("time_frame","percentage_difference",diff_early:diff_end) %>% 
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = percentage_difference,
      color = time_frame,
      fill = time_frame
    ),
    stat = "identity",
    position = "dodge"
  ) +
  coord_flip() +
  theme_classic()


ggsave(my_path("R","figures","rel_diff_per_species_mcp.png"),
       plot = rel_diff_mcp,
       width = 10,
       height = 8)

print(rel_diff_mcp)

  
```

This table shows the percentage difference between the MPA and the no-MPA runs

```{r per_diff_tablemcp, eval = T, message=F,warning=F}

print(mcp_relative_df)

```

## Per grid

### Abundance

#### Grid cell Status

Each grid cell is classified under a conservation status. Grid cells that have an MPA are classified as *protected*, those contiguous are *surrounding* and all other grids are *non-protected*. Interestingly, due to connectivity of grid cells, most-if-not-all of the Northern Shelf Bioregion is under some level of protection. 


```{r grid_status}

# Create gird status
status_grid <- my_path("D","Spatial","mpa_cpaw.csv", read = T) %>% 
  rowid_to_column("index") %>% 
      filter(index %in% bc_grid) %>% 
  mutate(
    status = ifelse(x1 == 1, "non-protected",
                    ifelse(x1 > 1, "surrounding","protected"))
  ) %>% 
  left_join(my_data("dbem_coords"))

# Save for further analysis
write_csv(status_grid,my_path("D","spatial","status_grid.csv"))

ggplot(status_grid) +
  geom_tile(aes(x =lon,y=lat,fill=status)) +
  geom_sf(data = bs_sf,aes(), fill ="transparent")

```


```{r abd_grid_df}

status_grid <- my_path("D","spatial","status_grid.csv",read = T)

variable = "Abd"

dbem_paths <- taxon_list %>% 
  distinct() %>% 
  select(taxon_key) %>% 
  mutate(dbem_no_mpa_path = paste0("~/Library/CloudStorage/OneDrive-UBC/Data/dbem/dbem_cmip6/r_data/C6GFDL26F1_t2/",taxon_key,variable,".RData"),
         dbem_mpa_path = paste0("~/Library/CloudStorage/OneDrive-UBC/Data/cpaws_bc/Results/raw/C6GFDL26F1CPAW/",taxon_key,variable,".RData")
  )


grid_per_species_abd <- agg_per_spp_abd(dbem_paths, grid = T)

write_csv(grid_per_species_abd, my_path("R","tables","grid_per_species_abd.csv"))

head(grid_per_species_abd)


grid_per_species_abd_scen <- bind_rows(
  lapply(scenarios, agg_all_scen, ssp = 26, grid = T)
  
)

```

#### Total abundance per grid cell status

Here we are simply looking at the total abundance sum by grid cell status for each species and time period. As before, it is expected that the MPA runs have more than the non-MPA runs, especially within the *protection* category. 


```{r grid_abd, fig.height=20, fig.width=15}

grid_per_species_abd <- my_path("R","tables","grid_per_species_abd.csv", read = T) %>% 
  left_join(taxon_list) %>% 
  filter(taxon_key != 601732) # remove mola mola


grid_abd_plot <- grid_per_species_abd %>% 
  filter(time_frame != "historic") %>% 
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = sum,
      color = run,
      fill = run
    ),
    stat = "identity",
    position = "dodge"
  ) +
  coord_flip() +
  facet_grid(status~time_frame,scales = "free") +
  theme_classic() +
  theme(
  text = element_text(size = 15)  # Set base text size
  )

ggsave(my_path("R","figures","grid_per_species_abd.png"),
       plot = grid_abd_plot,
       width = 15,
       height = 20)

print(grid_abd_plot)


```


#### Difference in abundance per grid cell

Figure showing the difference in abundance between the MPA and the no-MPA runs per species and per grid cell category. Percentage difference in abundance ($\Delta{abd}$) was calculated for each category as:

$$ \Delta{abd} = \frac{mpa-nompa}{nompa} * 100$$

Thus, results should be read as; positive values representing more abundance in the MPA runs and negative values more abundance in the non-MPA runs. Note that $ \Delta{abd}$ was caped at 100%/-100% for visualization purposes. One can argue that the larger the difference, the larger the effect of MPA in the population.

**Note:** This type of result has been contested previously on peer review as "logical" due to the nature of the model.

```{r grid_diff_abd, eval = T, fig.width=16, fig.height=20, message=F, warning = F}

# Raw data
grid_diff_plot <-
  grid_per_species_abd %>% 
  filter(time_frame != "historic") %>% 
  select(-variable,-mean) %>% 
  group_by(status, taxon_key,common_name,time_frame) %>%
  summarize(diff = sum[run == "mpa"] - sum[run == "no_mpa"],
            percent_diff = (diff / sum[run == "no_mpa"]) * 100
  ) %>% 
  mutate(percent_diff_capped = pmax(pmin(percent_diff, 100), -100)) %>% 
  # View()
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = percent_diff_capped,
      fill = percent_diff_capped
    ),
    stat = "identity",
    position = "dodge"
  ) +
  theme_classic()+
  coord_flip() +
  theme(
  text = element_text(size = 15)  # Set base text size
  ) +
  facet_grid(status~time_frame, scales = "free") +
  scale_fill_gradient2()


ggsave(my_path("R","figures","grid_diff_per_species_abd.png"),
       plot = grid_diff_plot,
       width = 15,
       height = 20)

print(grid_diff_plot)

```

### MCP 

```{r mcp_grid_df}

status_grid <- my_path("D","spatial","status_grid.csv",read = T)

variable = "Catch"

dbem_paths <- taxon_list %>% 
  distinct() %>% 
  select(taxon_key) %>% 
  mutate(dbem_no_mpa_path = paste0("~/Library/CloudStorage/OneDrive-UBC/Data/dbem/dbem_cmip6/r_data/C6GFDL26F1_t2/",taxon_key,"MCP.RData"),
         dbem_mpa_path = paste0("~/Library/CloudStorage/OneDrive-UBC/Data/cpaws_bc/Results/raw/C6GFDL26F1CPAW/",taxon_key,variable,".RData")
  )

grid_per_species_mcp <- agg_per_spp_mcp(dbem_paths, grid = T)

write_csv(grid_per_species_mcp, my_path("R","tables","grid_per_species_mcp.csv"))
```

#### Total MCP per grid cell status

Here we are simply looking at the total abundance sum by grid cell status for each species and time period. As before, it is expected that the MPA runs have less than the non-MPA runs, especially within the *protection* category since fishing effort is limited in these cells. 


```{r grid_mcp, fig.height=20, fig.width=15}

grid_per_species_mcp <- my_path("R","tables","grid_per_species_mcp.csv", read = T) %>% 
  left_join(taxon_list) %>% 
  filter(taxon_key != 601732) # remove mola mola


grid_mcp_plot <- grid_per_species_mcp %>% 
  filter(time_frame != "historic") %>% 
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = sum,
      color = run,
      fill = run
    ),
    stat = "identity",
    position = "dodge"
  ) +
  coord_flip() +
  facet_grid(status~time_frame,scales = "free") +
  theme_classic() +
  theme(
  text = element_text(size = 15)  # Set base text size
  )

ggsave(my_path("R","figures","grid_per_species_mcp.png"),
       plot = grid_mcp_plot,
       width = 15,
       height = 20)

print(grid_mcp_plot)


```

#### Difference in MCP per grid cell

```{r grid_diff_mcp, eval = T, fig.width=16, fig.height=20, message=F, warning = F}

# Raw data
grid_diff_plot_mcp <-
  grid_per_species_mcp %>% 
  filter(time_frame != "historic") %>% 
  select(-variable,-mean) %>% 
  group_by(status, taxon_key,common_name,time_frame) %>%
  summarize(diff = sum[run == "mpa"] - sum[run == "no_mpa"],
            percent_diff = (diff / sum[run == "no_mpa"]) * 100
  ) %>% 
  mutate(percent_diff_capped = pmax(pmin(percent_diff, 100), -100)) %>% 
  # View()
  ggplot() +
  geom_bar(
    aes(
      x = common_name,
      y = percent_diff_capped,
      fill = percent_diff_capped
    ),
    stat = "identity",
    position = "dodge"
  ) +
  theme_classic()+
  coord_flip() +
  theme(
  text = element_text(size = 15)  # Set base text size
  ) +
  facet_grid(status~time_frame, scales = "free") +
  scale_fill_gradient2()


ggsave(my_path("R","figures","grid_diff_per_species_mcp.png"),
       plot = grid_diff_plot_mcp,
       width = 15,
       height = 20)

print(grid_diff_plot_mcp)

```


# Salmon 

- One question that came up was why chinook salmon 'reacts' differently to the other salmons - and I wasn't sure what to say. Is the life history data different

Without diving into the data, overall salmonids are hard to model because of their andronomus life cycle and the reduced information we have of their ocean-stage habitat. That being said, Chinook spends the most amount of time in the ocean so, I believe is a result of the life history parameters that were used to model the distribution. Specifically, chinook salmon has the lowest *Intrinsic R* rate (0.114) and the largest *Linf* (153.3), they also have the shallower max depth (200 m) and the smallest VBonK (0.15).

```{r get_salmon_lif_history}

salmons <- dbem_species %>% 
  filter(str_detect(taxon_name,"Oncorhynchus"),
         str_detect(comon_name,"salmon"),
         common_name != "Masu salmon")

for(i in 1:5){
  
  file <- paste0(salmons$taxon_key[i],".txt")
  
  path_read <- my_path("G","Species/TaxonDataC/",file)
  
  partial <- read_delim(path_read,
                        delim = "\t", escape_double = FALSE, 
                        col_names = FALSE, trim_ws = TRUE) %>% 
    mutate(taxon_key = salmons$taxon_key[i]) %>% 
    slice(-1,-2)
  
  if(i == 1){
    final <- partial
  }else{
    final <- bind_rows(partial,final)
  }
  
}


# final %>% 
#   left_join(dbem_species) %>% 
#   spread(X1,X2) %>% 
#   select(-genus,-species) %>% 
#   write_csv("salmon_life_history.csv")
```


